---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```


## Load in the FAS Litters Data
  Janitor pacakge cleans up names of the variables
  
```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'Select' --Dplyer funcitons (part of tidyverse)
  operates on columns

Choose some columns and not others

```{r}
select(litters_df, group, gd0_weight)
```

specify range of columns you want to include using ':'

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

Specify columns want to lose
```{r}
select(litters_df, -litter_number)
```

Renaming Columns (and selecting columns want to keep)...
```{r}
select(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Just renaming - not selecting
```{r}
rename(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Helper functions 
  Used in specific context to accomplish something but wouldn't   out of that context 
  
  starts with, ends with, contains everything
  
  Select helpers
  
  selects all variables that start with "gd" 
```{r}
select(litters_df, starts_with("gd"))
```

Move selected variable to the beginning of view (stil keeps everything else)
```{r}
select(litters_df, litter_number, everything())
```

Relocate & not do anything else 
```{r}
relocate(litters_df, litter_number)
```

## 'filter' function
  intended to remove or keep rows that you're interested in for    analysis
  operates on rows (where as select operates on columns)
  - have to give it some condition
     either true or fasle (if true then it will be kept)
 
 looking at gestational weight on day zero only where < 22 
```{r}
filter(litters_df, gd0_weight < 22)
```

  gestational weight on day zerio where >= 22

```{r}
filter(litters_df, gd0_weight >= 22)
```

gestational age == 20 (USE 2 EQUAL SIGNS, asking a question, not assigning to value to this variable)

```{r}
filter(litters_df, gd_of_birth == 20)
```


see where this is true, and then flip
```{r}
filter(litters_df, !(gd_of_birth == 20))
```

does the same thing but ^ above is better for more complicated statements
```{r}
filter(litters_df, gd_of_birth != 20)
```


Want 2 things to be true
```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

Character variables
```{r}
filter(litters_df, group == "Con7")
```

```{r}
filter(litters_df, group == "Mod8")
```

2 character variables at the same time = to con 7 OR mod 8 
  %in% = where group is in either of these 2 options
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## 'mutate'
- modify existing variable
- create new variable 

Create variable 'weight_gain' by taking one variable subtracted by another 

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```

Creating & modifying variables at the same time
```{r}
    mutate(litters_df,
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
```


## 'arrange'
Goal is to put things in order
Primarily need to do if updating a data set to inspect in a way that's more intuitive than what's been put in
- shouldn't change analysis or resutls 
- helpful for knowning what's going on 
- alphabetical, numberical

order the pups born alive smallest to largest
```{r}
arrange(litters_df, pups_born_alive)
```

arrange 2 variables at the same time
```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```

