---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```


## Load in the FAS Litters Data
  Janitor pacakge cleans up names of the variables
  
```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'Select' --Dplyer funcitons (part of tidyverse)

Choose some columns and not others

```{r}
select(litters_df, group, gd0_weight)
```

specify range of columns you want to include using ':'

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

Specify columns want to lose
```{r}
select(litters_df, -litter_number)
```

Renaming Columns (and selecting columns want to keep)...
```{r}
select(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Just renaming - not selecting
```{r}
rename(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Helper functions 
  Used in specific context to accomplish something but wouldn't   out of that context 
  
  starts with, ends with, contains everything
  
  Select helpers
  
  selects all variables that start with "gd" 
```{r}
select(litters_df, starts_with("gd"))
```

Move selected variable to the beginning of view (stil keeps everything else)
```{r}
select(litters_df, litter_number, everything())
```

Relocate & not do anything else 
```{r}
relocate(litters_df, litter_number)
```


  


