---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Pivoting from wide format to long format
### pivot_longer
  Load the pulse data
  
    
```{r}
pulse_data = 
  haven:: read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Wide format to long format ...

   for each ID we have 4 observations on bdi    score and visit on which those things           happened
    - instead of 4 columns for bdi score
    - want for each ID, 4 rows with bdi scores         for each month   
    - names_prefix = "bdi_score_" removes that     prefix from the variable name to just get     bl, 01 m, 06, 12m

```{r}
pulse.data_tidy = 
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
  
```

Rewrite, combine, and extend (to add a mutate step)

```{r}
pulse_data = 
  haven:: read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id, visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
```

## 'pivot_wider'


Make up some data to show pivot wider

```{r}
analysis_result = 
  tibble(
      group = c("treatment", "treatment", "placebo", "placebo"), 
      time = c("pre", "post", "pre", "post"),
      mean = c (4, 8, 3.5, 4)
  )

analysis_result %>% 
    pivot_wider(
        names_from = "time", 
        values_from = "mean"
    )
```






